---
title: "ASICS package"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ASICS package}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

This vignette provides an overview of the package ASICS. ASICS is a completely 
automated procedure to identify and quantify metabolites in 1H NMR spectra of 
biological mixtures. It will enable empowering NMR-based metabolomics by quickly 
and accurately helping experts to obtain metabolic profiles.

```{r}
library(ASICS)
```

## Library of pure NMR spectra

A dataset with spectra of 175 pure metabolites is available.

Data are loaded and available metabolites are displayed with:
```{r lib}
head(pure_library$name, n = 20)
```

## Quantification of one or more Bruker files with ASICS

It is possible to perform the quantification with only one function directly 
from raw files (Bruker files):
```{r ASICS_onefile}
result <- ASICS(path = system.file("extdata", "example_spectra",
                                   "AG_faq_Beck01", package = "ASICS"),
                exclusion.areas = matrix(c(4.5,5.1,5.5,6.5),
                                         ncol = 2, byrow = TRUE))
print(result)
summary(result)
```

We can judge the quality of the quantification by stacking the original and the 
recomposed spectrum. Pure metabolites spectrum can also be added.
```{r plot_spectrum, warning=FALSE, fig.width=10, fig.height=6}
plot(result, xmin = 1, xmax = 1.5, ymax = 10, add_metab = "Lactate")
```

Relative concentrations of identified metabolites are saved in:
```{r rel_conc}
result@present_metabolites
```

Usually, many spectra are available for one experience. With the 
\code{ASICS_multiFiles} function, quantification is lauched on all spectra 
present in a folder (name.dir). Arguments are same as for \code{ASICS} function 
plus \code{ncores} for the parallel environment if wanted.

```{r ASICS_multifile, results = 'hide', eval = FALSE}
res_multi <- ASICS_multiFiles(name.dir = system.file("extdata",
                                                     "example_spectra",
                                                     package = "ASICS"),
                              exclusion.areas = matrix(c(4.5,5.1,5.5,6.5),
                                                       ncol = 2, byrow = TRUE),
                              ncores = 4)
```

For further analysis, it is possible to extract concentrations of all spectra in
a data-frame:
```{r extract_concentrations, eval = FALSE}
quantification <- extract_concentrations(res_multi)
```



