---
title: "ASICS"
author: "Gaëlle Lefort"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
package: ASICS
vignette: >
  %\VignetteIndexEntry{ASICS}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

This vignette provides an overview of the package ```ASICS```. ```ASICS``` is a 
fully automated procedure to identify and quantify metabolites in $^1$H NMR
spectra of biological mixtures. It will enable empowering NMR-based metabolomics
by quickly and accurately helping experts to obtain metabolic profiles.

```{r ASICSload}
library(ASICS)
```

## Library of pure NMR metabolites spectra

A dataset with spectra of 175 pure metabolites is available in the package. It
is automatically loaded at package start and available metabolites are displayed
with:
```{r lib}
head(pure_library$name, n = 20)
```

## Identification and quantification of metabolites on one or more Bruker files with ASICS

The identification and the quantification are carried out using only one 
function `ASICS` directly from raw files (Bruker files):
```{r ASICS_onefile}
# part of the spectrum to exclude (water)
to_exclude <- matrix(c(4.5,5.1,5.5,6.5), ncol = 2, byrow = TRUE)
result <- ASICS(path = system.file("extdata", "example_spectra",
                                   "AG_faq_Beck01", package = "ASICS"),
                exclusion.areas = to_exclude)
result
```

The quality of the results can be assessed by stacking the original and 
the recomposed spectrum on one plot. A pure metabolite spectrum can also be 
added for visual comparison.
```{r plot_spectrum, warning=FALSE, fig.width=8, fig.height=5}
plot(result, xmin = 1, xmax = 1.5, ymax = 10, add_metab = "Lactate")
```

Relative concentrations of identified metabolites are saved in:
```{r rel_conc}
present_metabolites(result)
```


## User's guide 

More details on ASICS functionnalities can be obtain in the user's guide :
```{r PDF, eval=FALSE}
ASICSUsersGuide()
```

## Reference 

Tardivel P., Canlet C., Lefort G., Tremblay-Franco M., Debrauwer L., Concordet 
D., Servien R. (2017). ASICS: an automatic method for identification and 
quantification of metabolites in complex 1D 1H NMR spectra. *Metabolomics*,
**13**(10): 109. https://doi.org/10.1007/s11306-017-1244-5


## Session information

This vignette has been created with the following system configuration
```{r sysinfo}
sessionInfo()
```




