---
title: "ASICS User's Guide"
author: "Gaëlle Lefort, Rémi Sevien, Patrick Tardivel and Nathalie Villa-Vialaneix"
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document2:
    toc_float: true
vignette: |
  %\VignetteIndexEntry{ASICS}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---


This user's guide provides an overview of the package `ASICS`. `ASICS` is a 
fully automated procedure to identify and quantify metabolites in $^1$H 1D-NMR 
spectra of biological mixtures. It will enable empowering NMR-based metabolomics
by quickly and accurately helping experts to obtain metabolic profiles. In 
addition to the quantification method, several functions allowing spectrum 
preprocessing or statistical analyses of quantified metabolites are available.

```{r ASICSload}
library(ASICS)
```

TODO: note sur le parallèle

# Context/Data

TODO: voir d'abord si toute l'analyse est faisable sans pb sur ce jeu de données.

Parler aussi de ASICS (reference à l'article). ici ?

# Library of pure NMR metabolite spectra

An object of class `PureLibrary` with spectra of 180 pure metabolites is 
required to perform the quantification. These spectra are metabolite spectra used
as a reference for quantification: only 
metabolites present in the library object can be quantified with ASICS.

A default library is available and is automatically loaded at package start. Available 
metabolites are displayed with:
```{r lib}
head(get_sample_name(pure_library), n = 8)
```

This library can be complemented or another library can be created with 
new spectra of pure metabolite. These spectra are imported from Bruker files and
a new library can be created with:
```{r import_create_lib, results='hide'}
pure_spectra <- import_spectra_bruker(system.file("extdata", "example_library", 
                                                  package = "ASICS"))
new_pure_library <- create_pure_library(pure_spectra, 
                                        nb.protons = c(5, 4, 9, 9))
```
A new library can also be created from txt or csv file, with samples in columns 
and chemical shifts in rows (see help page of `create_pure_library` function for
all details).


The newly created library can be used for quantification or merged with another one:
```{r select_merge_lib}
merged_pure_library <- c(pure_library[1:10], new_pure_library)
```
The `PureLibrary` `merged_pure_library` contains the first ten spectra of the
default library and the four newly imported spectra imported.



# Identification and quantification of metabolites on Bruker files with ASICS

First, data are imported in a data frame from Bruker files with the 
`import_spectra_bruker` function. These spectra are baseline corrected 
(Wang *et al*, 2013) and normalised by the area under the curve.
```{r import_from_Bruker, results='hide'}
spectra_data <- import_spectra_bruker(system.file("extdata", 
                                                  "Human_diabetes_example", 
                                                  package = "ASICS"))
```

Data can also be impored from other file types with standard `R` function. 
The only constraint is to have a data frame with spectra in columns 
(column names are sample names) and chemical shifts in rows (row names correspond to 
the ppm grid).
```{r import_from_txt, results='hide'}
diabetes <- system.file("extdata", "spectra_diabetes_example.txt", package = "ASICS")
spectra_data_txt <- read.table(diabetes, header = TRUE, row.names = 1)
```

Several functions for the preprocessing of spectra are also available: baseline 
correction, normalisation by the AUC and alignment (Vu et al., 2011).
```{r preprocessing, results='hide'}
spectra_base_cor <- baseline_correction(spectra_data_txt)
spectra_align <- alignment(spectra_base_cor)
spectra_norm <- normalisation(spectra_align)
```
Note than a normalisation by the area under the curve is included in the 
baseline correction and in the alignment functions. Moreover, baseline correction and
normalisation are also included in the `import_spectra_bruker` function.


Finally, from the data frame, a `Spectra` object is created. This is a required 
step for the quantification.
```{r create_spectra, results='hide'}
spectra_obj <- create_spectra(spectra_align)
```

--- ok jusqu'ici


Identification and quantification of metabolites can now be carried out using only 
the function `ASICS`. This function takes approximately 2 minutes by 
spectrum to run. A parallel implementation is available for unix-like OS, with the number of 
CPU cores on the current host - 2 is created by default. Parallel environment
for spectra importation and quantification estimation is switched on/off with
the option `parallel`.
```{r ASICS, results='hide', cache = TRUE, dependson='import_from_Bruker'}
# part of the spectrum to exclude (water and urea)
to_exclude <- matrix(c(4.5, 5.1, 5.5, 6.5), ncol = 2, byrow = TRUE)
system.time(ASICS_results <- ASICS(spectra_obj[1], exclusion.areas = to_exclude))
```

Summary of ASICS results:
```{r summary_res}
ASICS_results
```


The quality of the results can be assessed by stacking the original and 
the reconstructed spectra on one plot. A pure metabolite spectrum can also be 
added for visual comparison. For example, the first spectrum with Lactate:
```{r plot_spectrum, warning=FALSE, fig.width=12, fig.height=8}
plot_spectrum(ASICS_results, idx = 1, xlim = c(1, 1.5), add.metab = "Lactate")
```

Relative concentrations of identified metabolites are saved in a data frame
accessible via the `get_quantification` function:
```{r rel_conc}
head(get_quantification(ASICS_results), 10)[, 1:2]
```

# Analysis on quantification (PCA, OPLS/OPLS-DA and tests)



```{r}
design <- read.table(system.file("extdata", "design_diabete_example.txt", 
                                 package = "ASICS"), header = TRUE)
```


```{r}
analysis_data <- format_for_analysis(get_quantification(ASICS_results),
                                     design = design, zero.threshold = 0.5, 
                                     outliers = c("ADG10003u_007", "ADG19007u_163"))
```


PCA : 
```{r, fig.width=10}
resPCA <- PCA(analysis_data)
plot_PCA(resPCA, graph = "ind", col.ind = "condition")
plot_PCA(resPCA, graph = "var")
```


OPLS-DA:
```{r}
resOPLSDA <- OPLSDA(analysis_data, condition = "condition")
plot_OPLSDA(resOPLSDA)
```

Tests:
```{r}
kruskal_on_quantification(analysis_data, "condition")
```




# References 


Tardivel P., Canlet C., Lefort G., Tremblay-Franco M., Debrauwer L., Concordet 
D., Servien R. (2017). ASICS: an automatic method for identification and 
quantification of metabolites in complex 1D 1H NMR spectra. *Metabolomics*,
**13**(10): 109. https://doi.org/10.1007/s11306-017-1244-5

Wang, K. C., Wang, S. Y., Kuo, C. H., Tseng, Y. J. (2013). Distribution-based 
classification method for baseline correction of metabolomic 1D proton nuclear 
magnetic resonance spectra. *Analytical Chemistry*, **85**(2), 1231–1239.

Vu, T. N., Valkenborg, D., Smets, K., Verwaest, K. A., Dommisse, R., Lemiere, 
F., ... & Laukens, K. (2011). An integrated workflow for robust alignment and 
simplified quantitative analysis of NMR spectrometry data. 
*BMC bioinformatics*, **12**(1), 405.

# Session information

This user's guide has been created with the following system configuration
```{r sysinfo}
sessionInfo()
```

