---
title: "ASICS User's Guide"
author: "Gaëlle Lefort"
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document2:
    toc_float: true
vignette: |
  %\VignetteIndexEntry{ASICS}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---


This user's guide provides an overview of the package `ASICS`. `ASICS` is a 
fully automated procedure to identify and quantify metabolites in $^1$H NMR 
spectra of biological mixtures. It will enable empowering NMR-based metabolomics
by quickly and accurately helping experts to obtain metabolic profiles.

```{r ASICSload}
library(ASICS)
```

# Data

TODO: recopier article

# Library of pure NMR metabolite spectra

An object of class `PureLibrary` with spectra of 174 pure metabolites is 
available in the package. It is automatically loaded at package start and 
available metabolites are displayed with:
```{r lib}
head(get_sample_name(pure_library), n = 8)
```

Pure spectra can be imported from Bruker files and a new library can be created
with:
```{r import_create_lib, results='hide'}
pure_spectra <- import_spectra_bruker(system.file("extdata", "example_library", 
                                                  package = "ASICS"))
new_pure_library <- create_pure_library(pure_spectra, 
                                        nb.protons = c(5, 4, 9, 9))
```

New library can be used for quantification or merge with another one:
```{r select_merge_lib}
merged_pure_library <- c(pure_library[1:10], new_pure_library)
```
The `PureLibrary` `merged_pure_library` contains the first ten spectra of 
default library and the four spectra imported above.



# Identification and quantification of metabolites on Bruker files with ASICS

First, data are imported in a data frame from Bruker files with the 
`import_spectra_bruker` function. These spectra are baseline corrected 
(Wang *et al*, 2013) and normalised by the area under the curve.
It also possible to import data already cleaned from other file types with usual 
`R` function. The only constraint is to have a data frame with spectra in colums 
(colnames are sample names) and chemical shifts in rows (rownames correspond to 
the ppm grid).
```{r import_from_Bruker, results='hide'}
spectra_data <- import_spectra_bruker(system.file("extdata", 
                                                  "Human_diabetes_example", 
                                                  package = "ASICS"))
```

Then, from the data frame, a `Spectra` object is created. This is a required 
step for the quantification.
```{r create_spectra, results='hide'}
spectra_obj <- create_spectra(spectra_data)
```

It is now possible to carry out the identification and quantification using only 
the function `ASICS`. This function can take some times (about 1min30 by 
spectrum) thus, execpt on Windows, a parallel environment with the number of 
CPU cores on the current host - 2 is created by default. This also applies for 
data import. If you don't want a parallel environment, you can set `parallel` 
parameter to `FALSE`.
```{r ASICS, results='hide', cache = TRUE, dependson='import_from_Bruker'}
# part of the spectrum to exclude (water and urea)
to_exclude <- matrix(c(4.5, 5.1, 5.5, 6.5), ncol = 2, byrow = TRUE)
system.time(ASICS_results <- ASICS(spectra_obj[1], exclusion.areas = to_exclude))
```

Summary of ASICS results:
```{r summary_res}
ASICS_results
```


The quality of the results can be assessed by stacking the original and 
the recomposed spectra on one plot. A pure metabolite spectrum can also be 
added for visual comparison. For example, the first spectrum with Lactate:
```{r plot_spectrum, warning=FALSE, fig.width=12, fig.height=8}
plot_spectrum(ASICS_results, idx = 1, xlim = c(1, 1.5), add.metab = "Lactate")
```

Relative concentrations of identified metabolites are saved in a data frame
accessible via the `get_quantification` function:
```{r rel_conc}
head(get_quantification(ASICS_results), 10)[, 1:2]
```

# Analysis on quantification (PCA, OPLS/OPLS-DA and tests)



```{r}
design <- read.table(system.file("extdata", "design_diabete_example.txt", 
                                 package = "ASICS"), header = TRUE)
```


```{r}
analysis_data <- format_for_analysis(get_quantification(ASICS_results),
                                     design = design, zero.threshold = 0.5, 
                                     outliers = c("ADG10003u_007", "ADG19007u_163"))
```


PCA : 
```{r, fig.width=10}
resPCA <- PCA(analysis_data)
plot_PCA(resPCA, graph = "ind", col.ind = "condition")
plot_PCA(resPCA, graph = "var")
```


OPLS-DA:
```{r}
resOPLSDA <- OPLSDA(analysis_data, condition = "condition")
plot_OPLSDA(resOPLSDA)
```

Tests:
```{r}
kruskal_on_quantification(analysis_data, "condition")
```




# References 


Tardivel P., Canlet C., Lefort G., Tremblay-Franco M., Debrauwer L., Concordet 
D., Servien R. (2017). ASICS: an automatic method for identification and 
quantification of metabolites in complex 1D 1H NMR spectra. *Metabolomics*,
**13**(10): 109. https://doi.org/10.1007/s11306-017-1244-5

Wang, K. C., Wang, S. Y., Kuo, C. H., Tseng, Y. J. (2013). Distribution-based 
classification method for baseline correction of metabolomic 1D proton nuclear 
magnetic resonance spectra. *Analytical Chemistry*, **85**(2), 1231–1239.


# Session information

This user's guide has been created with the following system configuration
```{r sysinfo}
sessionInfo()
```

